The Unified Nature of the Dielectric Properties of Two-Dimensional Materials

### 2018-11-15 SI.tex
Correct minor grammatic errors

\author{Tian Tian}
\affiliation{Institute for Chemical and Bioengineering, ETH Z{\"{u}}rich,  Vladimir Prelog Weg 1, CH-8093 Z{\"{u}}rich, Switzerland}
\altaffiliation{T. T. and D. S. contributed equally to this work}
\author{Declan Scullion}
\affiliation{School of Mathematics and Physics, Queen's University Belfast, BT7 1NN, United Kingdom}
\altaffiliation{T. T. and D. S. contributed equally to this work}
\author{Dale Hughes}
\affiliation{School of Mathematics and Physics, Queen's University Belfast, BT7 1NN, United Kingdom}
\author{Lu Hua Li}
\affiliation{Institute for Frontier Materials, Deakin University, Waurn Ponds, Victoria, Australia}
\author{Chih-Jen Shih}
\affiliation{Institute for Chemical and Bioengineering, ETH Z{\"{u}}rich,  Vladimir Prelog Weg 1, CH-8093 Z{\"{u}}rich, Switzerland}
\author{Jonathan Coleman}
\affiliation{School of Physics, Centre for Research on Adaptive Nanostructures and Nanodevices (CRANN) and Advanced Materials and BioEngineering Research (AMBER), Trinity College Dublin, Dublin 2, Ireland.}
\author{Manish Chhowalla}
% \affiliation{Materials Science and Engineering, Rutgers University, 607 Taylor Road, Piscataway, New Jersey 08854, USA.}
\affiliation{Department of Materials Science \& Metallurgy, University of Cambridge, CB3 0FS, United Kindom}
\author{Elton J. G. Santos}
\email{e.santos@qub.ac.uk}
\affiliation{School of Mathematics and Physics, Queen's University Belfast, BT7 1NN, United Kingdom}

@@

\title{Universal Understanding of the Dielectric Nature of Two-Dimensional Materials}
\begin{document}

@@

To conclude, based on theoretical and technical considerations, there are several
advantages of using the polarizability $\alpha$ for describing the dielectric
nature of 2D materials, including
\begin{enumerate}
\item $\alpha_{\mathrm{2D}}$ can be used to describe both the local and macroscopic dielectric properties, while $\epsilon_{\mathrm{2D}}$ can't.
\item Calculating $\alpha_{\mathrm{2D}}$ only requires to calculate the dielectric response at single superlattice length, while $\epsilon_{\mathrm{2D}}$ requires calculation with varied superlattice length.
\item $\alpha_{\mathrm{2D}}$ correctly respresents the screening length of 2D material, while $\epsilon_{\mathrm{2D}}$ does not.
\item $\alpha_{\mathrm{2D}}$ correctly represents the different degree of screening in/out-of-plane, while $\epsilon_{\mathrm{2D}}$ does not.

@@

We also investigate the relation of 2D polarizabilities with
difference choices of \textit{ab initio} bandgaps. It is widely
accepted that the PBE exchange correlation, tends to underestimate the
bangap \cite{Heyd_2005,Kumar_2016_PRB,Kumar_2016_jpcc}.  Indeed,
changing the choice of $E_{\mathrm{g}}$ yields different regression
relation with $1/\alpha_{\mathrm{2D}}^{\parallel}$, as shown in Figure
\ref{fig:alpha-Eg-diff}. We see that due to the underestimation of PBE
bandgap, the slope of linear regression is larger than that from
HSE-bandgap. We also observe that the $1/\alpha-E_{\mathrm{g}}$
relation is better presented by using the minimal HSE bandga than the
minimal PBE bandgap, due to higher regression $R^{2}$ coefficient of
the former. We note that the higher $R^{2}$ coefficient observed using
the direct PBE bandgap than the minimal PBE bandgap may be solely
caused by the fact that the direct bandgap of 2D materials on the PBE
level is closer to the HSE bandgap. From the random phase
approximation theory of dielectric response, the polarizability is
contributed by all possible transition between valence and conduction
bands, with the minimal bandgap as the least possible transition. In
this sense, $\alpha_{\mathrm{2D}}^{\parallel}$ is mostly like to be associated with
the minimal, not direct bandgap, as also observed in the original Moss
relation. We also examine the validity of such statement based on the
analysis of the Computational 2D Materials Database (C2DB)
\cite{Haastrup_2018}, as will be discussed in the following sections.

@@

Due to the high computation complexity of dielectric response
calculations on the HSE level, our 2D material database is limited to
about 50 types of materials. It is desirable to validate our proposed
relations on a even larger scale database. We select over 230
semiconducting 2D materials from the C2DB database with a GW bandgap
larger than 0.05 eV and extracted the 2D polarizabilities calculated
on the PBE level. The propsed linear relations between
$1/\alpha_{\mathrm{2D}}^{\parallel}-E_{\mathrm{g}}(\mathrm{HSE})$ and
$\alpha_{\mathrm{2D}}^{\perp}-\delta_{\mathrm{cov}}$ are also valid,
as shown in Figure \ref{fig:gpaw-alpha-relation}. Excellent linear
correlation is observed in both cases, with the $R^{2}$ coefficient
larger than 0.9, indicating the existence of the universal description
of 2D dielectric nature through the proposed relations with the 2D
polarizabilities. We note that the slope of linear regression is
slightly different from the our proposed relation from the dielectric
response on HSE06 level, due to the fact that the macroscopic
dielectric constant $\varepsilon_{\mathrm{SL}}$ calculated on the PBE
level is about slightly difference. Despite the minor deviation of the
slope of linear regression, the analysis on the C2DB database supports
our understanding of the universal dielectric nature of 2D materials.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\linewidth]{img/SI-gpaw-alpha_Eg_all.pdf}
  \caption{$\alpha_{\mathrm{2D}}^{\parallel}$ as funtion of minimal and direct
    $E_{\mathrm{g}}$ calculated on different theoretical levels:
    \textbf{a} PBE, \textbf{b} HSE and \textbf{c} GW of the C2DB
    database.}
  \label{fig:SI-gpaw-alpha-Eg-all}
\end{figure}

For 2D materials, the exciton effect plays an important role in
determining the experimentally accessible bandgap
\cite{Arnaud_2006_exc_hBN,Pulci_2014,Ramasubramaniam_2012,Chernikov_2014_EB_MoS2_2D3D}. The
experimentally observed optical bandgap
$E_{\mathrm{g}}^{\mathrm{opt}}$, is usually lower than the direct
bandgap from bandstructure $E_{\mathrm{g}}^{\mathrm{dir}}$ by the
exciton binding energy $E_{\mathrm{b}}$, which is at the 10$^{-1}$ to
10$^{1}$ eV for different 2D materials due to the attenuated
dielectric screening. Next we examine the relation between
$\alpha_{\mathrm{2D}}^{\parallel}$ and $E_{\mathrm{g}}^{\mathrm{opt}}$
from the C2DB database, with
$E_{\mathrm{g}}^{\mathrm{opt}}=E_{\mathrm{g}}^{\mathrm{dir,QP}}-E_{\mathrm{b}}^{\mathrm{BSE}}$,
where $E_{\mathrm{g}}^{\mathrm{dir,QP}}$ is the direct quasi-particle
bandgap calculated using G0W0 method and
$E_{\mathrm{b}}^{\mathrm{BSE}}$ is the exciton binding energy
calculated using the Bethe-Salpethe equation. Figure \ref{fig:opt}
shows $(4\pi\varepsilon_{0})/\alpha_{\mathrm{2D}}^{\parallel}$ as a
function of $E_{\mathrm{g}}^{\mathrm{opt}}$, with a linear regression
slope of 0.154 and $R^{2}$ of 0.84, similar to the relation between
$(4\pi \varepsilon_{0})/\alpha_{\mathrm{2D}}^{\parallel}$ and
$E_{\mathrm{g}}$ (from HSE06 level, see Figures \ref{fig:SI-raw-HSE}
and \ref{fig:SI-gpaw-alpha-Eg-all}). The roughly linear correlation
between $(4\pi \varepsilon_{0})/\alpha_{\mathrm{2D}}^{\parallel}$ and
$E_{\mathrm{g}}^{\mathrm{opt}}$ is not coincidence: in fact,
theoretical analysis shows that the binding energy $E_{\mathrm{b}}$ is
proportional to the direct bandgap $E_{\mathrm{g}}$
\cite{Jiang_2017_Eg_Eb}, taking into account that
$(\alpha_{\mathrm{2D}}^{\parallel})^{-1} \propto E_{\mathrm{g}}$, we
rationalize that
$(\alpha_{\mathrm{2D}}^{\parallel})^{-1} \propto
E_{\mathrm{g}}^{\mathrm{opt}}=E_{\mathrm{g}}^{\mathrm{dir}}-E_{\mathrm{b}}$. The
slightly smaller linearity than the 2D Moss-like relation is caused
from multiple approximations used. Nevertheless we show that
$(4\pi\varepsilon_{0})/\alpha_{\mathrm{2D}}^{\parallel}$ can be
equivalently predicted using the experimentally accessible optical
bandgap.

@@

In this section we show in detail the polarizability-based theoretical
framework that leads to the 2D Moss-like relations proposed in the
main text.  Due to its highly anisotropic nature, the wavefunction of
an isolated 2D material $\psi(\mathbf{r})$ can be separated into the
in- and out-of-plane components, similar to the treatment of quantum
wells (QW),\cite{davies_physics_1997} such that
$\psi^{\parallel}(\boldsymbol{\rho})$ and $\psi^{\perp}(z)$, such that
$\psi=\psi^{\parallel}\psi^{\perp}$, where $\boldsymbol{\rho}=(x, y)$
is the in-plane coordinate. Using the Bloch theorem, the periodic
$\psi^{\parallel}(\boldsymbol{\rho})$ can be further expressed as
$\psi^{\parallel}(\boldsymbol{\rho})=e^{i\mathbf{k} \cdot
  \boldsymbol{\rho}}u(\boldsymbol{\rho})$, where $\mathbf{k}$ is the
in-plane wave vector and $u(\boldsymbol{\rho})$ is periodic function
in the xy-plane. According to the random phase approximation (RPA)
theory\cite{Adler_1962}, $\varepsilon_{\mathrm{SL}}$ is the
$\mathbf{q} \to 0$ and $\omega \to 0$ limits of the non-interacting
dielectric function $\varepsilon(\mathbf{q}, \omega)$, where
$\mathbf{q}$ is the momentum transfer and $\omega$ is the frequency:
\begin{equation}
  \label{eq:RPA-eps2}
  \varepsilon_{\mathrm{SL}}
  = \lim_{\mathbf{q} \to 0} 1 + \frac{2e^{2}}{\varepsilon_{0} |\mathbf{q}|^{2} \Omega}
  \sum_{\mathrm{k, c, v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
  \left[f(\psi_{\mathrm{c}}) - f(\psi_{\mathrm{v}})\right]
\end{equation}
where $e$ is the unit charge, c, v are the conduction and valence
bands, $E$ is the eigenenergy of individual bands, and $f$ is the
Fermi-Dirac distribution function. Taking the
limit that $L\to\infty$, when
$\varepsilon^{\perp}_{\mathrm{SL}} \approx 1$, we have
$1-1/\varepsilon^{\perp}_{\mathrm{SL}} \approx
(\varepsilon_{\mathrm{SL}}^{\perp} - 1)$. Therefore
$\alpha_{\mathrm{2D}}^{\parallel}$ and $\alpha_{\mathrm{2D}}^{\perp}$
at 0 K can be unified by the same equation:
\begin{equation}
  \label{eq:alpha-RPA}
  \alpha = \frac{2e^{2}}{|q|^{2}A} \sum_{\mathrm{k,c,v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
where the direction is determined by $\mathbf{q}$. Next we will show
that the different behavior of $\psi^{\parallel}$ and $\psi^{\perp}$
give rise to the main text Eqs. \ref{main-eq:2D-Moss-para} and
\ref{main-eq:2D-Moss-perp}.

@@

\subsubsection{Detailed derivations for main text Eq. \ref{main-eq:2D-Moss-perp}}
\label{ssec:theory-2}
For main text Eq. \ref{main-eq:2D-Moss-perp}, treating the in-plane
wavefunctions as plane wave with form
$\psi^{\parallel}(\rho) \propto e^{i \mathbf{k \rho}}$, the matrix
element of
$<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>$,
when $\mathbf{q}=(0, 0, q_{z})$, becomes\cite{Hybertsen_1987}:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z}
  <\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>
  &= \frac{1}{A} \int dx \int dy
  e^{i(\mathbf{-k \rho} - \mathbf{q \rho} + \mathbf{(k+q) \rho})}
  \int (\psi^{\perp})^{*}_{\mathrm{v}}(\mathbf{k})e^{-iq_{z}z}\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})\\
  &= <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})>
  \end{aligned}
\end{equation}
Note that the states perpendicular are bound, the integral is
meaningful only when $\mathbf{k=k+q}$ \cite{davies_physics_1997}. By
performing the Taylor expansion of
$e^{-i\mathbf{qr}} \approx 1 - i\mathbf{qr}$, we get:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z-2}
    <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
    &\approx <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})> -
    iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>\\
    &= -iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
   \end{aligned}
\end{equation}
plug this into Eq. \ref{eq:alpha-RPA} and express the summation over
$k_{x}$ and $k_{y}$ in a continuous form within the Brillouin Zone, we
arrive at:
\begin{equation}
\label{eq:alpha_perp_RPA}
\alpha_{\mathrm{2D}}^{\perp} = \frac{2e^{2}}{(2 \pi) ^{2}} \int d^{2}\mathbf{k}
\sum_{\mathrm{c, v}}
\frac{|<\psi_{\mathrm{v}}(\mathbf{k})|z|\psi_{\mathrm{c}}(\mathbf{k})>|^{2}}
{E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
The formalism is slightly different from Eq.\ref{eq:alpha_para_RPA}.

The out-of-plane wavefunction $\psi^{\perp}(z)$ is the solution to the Schr\"{o}dinger equation with
Hamiltonian $\mathcal{H} = -\hbar^{2} \nabla^{2}/2m_{e} + V(z)$, where
$\hbar$ is the reduced Planck constant, $m_{e}$ is electron mass and
$V(z)$ is the confined Coulombic potential along the z-direction
created by the nuclei
\cite{davies_physics_1997,ihn_semiconductor_2009}. Although the exact
form for $\psi^{\perp}$ depends on the exact distribution of $V(z)$,
without loss of generality we can assume the electrons are confined in
a potential well of width $\delta$, which is the typical treatment for
semiconductor QWs \cite{ihn_semiconductor_2009}. The allowed bound
states inside the confined region generally has wave vector
$k_{z} \propto n \pi / \delta$. With the total energy
$E_{n}(\mathbf{k}) = {\displaystyle \frac{\hbar^{2} (k_{x}^{2} +
    k_{y}^{2})}{2 m^{\parallel}} + \frac{\hbar^{2} n^{2} \pi^{2}}{2
    m^{\perp} \delta^{2}}}$, where $m^{\parallel}$ and $m^{\perp}$ are
the effective masses parallel and perpendicular to the 2D
plane. Therefore the denominator of Eq. 10 becomes independent of
$\mathbf{k}$, that
$E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k}) =
(n_{\mathrm{c}}^{2} - n_{\mathrm{v}}^{2}) {\displaystyle
  \frac{\hbar^{2} \pi^{2}}{2 m^{\perp} \delta^{2}}}$. On the other
hand, the numerator
$<\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>$
is proportional to the confinement length $\delta$ (can be seen using
particle-in-box solution\cite{davies_physics_1997}). In combination,
the individual terms of the summation in the right hand of Eq. 5b is
independent of neither $E_{\mathrm{g}}$ nor \textbf{k}, proving that
$\alpha_{\mathrm{2D}}^{\perp}$ is independent of the band gap.

>>>>

\author{Tian Tian}
\affiliation{Institute for Chemical and Bioengineering, ETH Z{\"{u}}rich,  Vladimir Prelog Weg 1, CH-8093 Z{\"{u}}rich, Switzerland}
\altaffiliation{T. T. and D. S. contributed equally to this work}
\author{Declan Scullion}
\affiliation{School of Mathematics and Physics, Queen's University Belfast, BT7 1NN, United Kingdom}
\altaffiliation{T. T. and D. S. contributed equally to this work}
\author{Dale Hughes}
\affiliation{School of Mathematics and Physics, Queen's University Belfast, BT7 1NN, United Kingdom}
\author{Lu Hua Li}
\affiliation{Institute for Frontier Materials, Deakin University, Waurn Ponds, Victoria, Australia}
\author{Chih-Jen Shih}
\affiliation{Institute for Chemical and Bioengineering, ETH Z{\"{u}}rich,  Vladimir Prelog Weg 1, CH-8093 Z{\"{u}}rich, Switzerland}
\author{Jonathan Coleman}
\affiliation{School of Physics, Centre for Research on Adaptive Nanostructures and Nanodevices (CRANN) and Advanced Materials and BioEngineering Research (AMBER), Trinity College Dublin, Dublin 2, Ireland.}
\author{Manish Chhowalla}
% \affiliation{Materials Science and Engineering, Rutgers University, 607 Taylor Road, Piscataway, New Jersey 08854, USA.}
\affiliation{Department of Materials Science \& Metallurgy, University of Cambridge, CB3 0FS, United Kingdom}
\author{Elton J. G. Santos}
\email{e.santos@qub.ac.uk}
\affiliation{School of Mathematics and Physics, Queen's University Belfast, BT7 1NN, United Kingdom}

@@

\title{The Unified Nature of the Dielectric Properties of Two-Dimensional Materials}
\begin{document}

@@

To conclude, based on theoretical and technical considerations, there are several
advantages of using the polarizability $\alpha$ for describing the dielectric
nature of 2D materials, including
\begin{enumerate}
\item $\alpha_{\mathrm{2D}}$ can be used to describe both the local and macroscopic dielectric properties, while $\epsilon_{\mathrm{2D}}$ can't.
\item Calculating $\alpha_{\mathrm{2D}}$ only requires to calculate the dielectric response at single superlattice length, while $\epsilon_{\mathrm{2D}}$ requires calculation with varied superlattice length.
\item $\alpha_{\mathrm{2D}}$ correctly represents the screening length of 2D material, while $\epsilon_{\mathrm{2D}}$ does not.
\item $\alpha_{\mathrm{2D}}$ correctly represents the different degree of screening in/out-of-plane, while $\epsilon_{\mathrm{2D}}$ does not.

@@

We also investigate the relation of 2D polarizabilities with
difference choices of \textit{ab initio} bandgaps. It is widely
accepted that the PBE exchange correlation, tends to underestimate the
bangap \cite{Heyd_2005,Kumar_2016_PRB,Kumar_2016_jpcc}.  Indeed,
changing the choice of $E_{\mathrm{g}}$ yields different regression
relation with $1/\alpha_{\mathrm{2D}}^{\parallel}$, as shown in Figure
\ref{fig:alpha-Eg-diff}. We see that due to the underestimation of PBE
bandgap, the slope of linear regression is larger than that from
HSE-bandgap. We also observe that the $1/\alpha-E_{\mathrm{g}}$
relation is better presented by using the minimal HSE bandgap than the
minimal PBE bandgap, due to higher regression $R^{2}$ coefficient of
the former. We note that the higher $R^{2}$ coefficient observed using
the direct PBE bandgap than the minimal PBE bandgap may be solely
caused by the fact that the direct bandgap of 2D materials on the PBE
level is closer to the HSE bandgap. From the random phase
approximation theory of dielectric response, the polarizability is
contributed by all possible transition between valence and conduction
bands, with the minimal bandgap as the least possible transition. In
this sense, $\alpha_{\mathrm{2D}}^{\parallel}$ is mostly like to be associated with
the minimal, not direct bandgap, as also observed in the original Moss
relation. We also examine the validity of such statement based on the
analysis of the Computational 2D Materials Database (C2DB)
\cite{Haastrup_2018}, as will be discussed in the following sections.

@@

Due to the high computation complexity of dielectric response
calculations on the HSE level, our 2D material database is limited to
about 50 types of materials. It is desirable to validate our proposed
relations on an even larger scale database. We select over 230
semiconducting 2D materials from the C2DB database with a GW bandgap
larger than 0.05 eV and extracted the 2D polarizabilities calculated
on the PBE level. The proposed linear relations between
$1/\alpha_{\mathrm{2D}}^{\parallel}-E_{\mathrm{g}}(\mathrm{HSE})$ and
$\alpha_{\mathrm{2D}}^{\perp}-\delta_{\mathrm{cov}}$ are also valid,
as shown in Figure \ref{fig:gpaw-alpha-relation}. Excellent linear
correlation is observed in both cases, with the $R^{2}$ coefficient
larger than 0.9, indicating the existence of the universal description
of 2D dielectric nature through the proposed relations with the 2D
polarizabilities. We note that the slope of linear regression is
slightly different from the our proposed relation from the dielectric
response on HSE06 level, due to the fact that the macroscopic
dielectric constant $\varepsilon_{\mathrm{SL}}$ calculated on the PBE
level is about slightly difference. Despite the minor deviation of the
slope of linear regression, the analysis on the C2DB database supports
our understanding of the universal dielectric nature of 2D materials.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\linewidth]{img/SI-gpaw-alpha_Eg_all.pdf}
  \caption{$\alpha_{\mathrm{2D}}^{\parallel}$ as function of minimal and direct
    $E_{\mathrm{g}}$ calculated on different theoretical levels:
    \textbf{a} PBE, \textbf{b} HSE and \textbf{c} GW of the C2DB
    database.}
  \label{fig:SI-gpaw-alpha-Eg-all}
\end{figure}

For 2D materials, the exciton effect plays an important role in
determining the experimentally accessible bandgap
\cite{Arnaud_2006_exc_hBN,Pulci_2014,Ramasubramaniam_2012,Chernikov_2014_EB_MoS2_2D3D}. The
experimentally observed optical bandgap
$E_{\mathrm{g}}^{\mathrm{opt}}$, is usually lower than the direct
bandgap from band structure $E_{\mathrm{g}}^{\mathrm{dir}}$ by the
exciton binding energy $E_{\mathrm{b}}$, which is at the 10$^{-1}$ to
10$^{1}$ eV for different 2D materials due to the attenuated
dielectric screening. Next we examine the relation between
$\alpha_{\mathrm{2D}}^{\parallel}$ and $E_{\mathrm{g}}^{\mathrm{opt}}$
from the C2DB database, with
$E_{\mathrm{g}}^{\mathrm{opt}}=E_{\mathrm{g}}^{\mathrm{dir,QP}}-E_{\mathrm{b}}^{\mathrm{BSE}}$,
where $E_{\mathrm{g}}^{\mathrm{dir,QP}}$ is the direct quasi-particle
bandgap calculated using G0W0 method and
$E_{\mathrm{b}}^{\mathrm{BSE}}$ is the exciton binding energy
calculated using the Bethe-Salpethe equation. Figure \ref{fig:opt}
shows $(4\pi\varepsilon_{0})/\alpha_{\mathrm{2D}}^{\parallel}$ as a
function of $E_{\mathrm{g}}^{\mathrm{opt}}$, with a linear regression
slope of 0.154 and $R^{2}$ of 0.84, similar to the relation between
$(4\pi \varepsilon_{0})/\alpha_{\mathrm{2D}}^{\parallel}$ and
$E_{\mathrm{g}}$ (from HSE06 level, see Figures \ref{fig:SI-raw-HSE}
and \ref{fig:SI-gpaw-alpha-Eg-all}). The roughly linear correlation
between $(4\pi \varepsilon_{0})/\alpha_{\mathrm{2D}}^{\parallel}$ and
$E_{\mathrm{g}}^{\mathrm{opt}}$ is not coincidence: in fact,
theoretical analysis shows that the binding energy $E_{\mathrm{b}}$ is
proportional to the direct bandgap $E_{\mathrm{g}}$
\cite{Jiang_2017_Eg_Eb}, taking into account that
$(\alpha_{\mathrm{2D}}^{\parallel})^{-1} \propto E_{\mathrm{g}}$, we
rationalize that
$(\alpha_{\mathrm{2D}}^{\parallel})^{-1} \propto
E_{\mathrm{g}}^{\mathrm{opt}}=E_{\mathrm{g}}^{\mathrm{dir}}-E_{\mathrm{b}}$. The
slightly smaller linearity than the 2D Moss-like relation is caused
from multiple approximations used. Nevertheless we show that
$(4\pi\varepsilon_{0})/\alpha_{\mathrm{2D}}^{\parallel}$ can be
equivalently predicted using the experimentally accessible optical
bandgap.

@@

In this section we show in detail the polarizability-based theoretical
framework that leads to the 2D Moss-like relations proposed in the
main text.  Due to its highly anisotropic nature, the wave function of
an isolated 2D material $\psi(\mathbf{r})$ can be separated into the
in- and out-of-plane components, similar to the treatment of quantum
wells (QW),\cite{davies_physics_1997} such that
$\psi^{\parallel}(\boldsymbol{\rho})$ and $\psi^{\perp}(z)$, such that
$\psi=\psi^{\parallel}\psi^{\perp}$, where $\boldsymbol{\rho}=(x, y)$
is the in-plane coordinate. Using the Bloch theorem, the periodic
$\psi^{\parallel}(\boldsymbol{\rho})$ can be further expressed as
$\psi^{\parallel}(\boldsymbol{\rho})=e^{i\mathbf{k} \cdot
  \boldsymbol{\rho}}u(\boldsymbol{\rho})$, where $\mathbf{k}$ is the
in-plane wave vector and $u(\boldsymbol{\rho})$ is periodic function
in the xy-plane. According to the random phase approximation (RPA)
theory\cite{Adler_1962}, $\varepsilon_{\mathrm{SL}}$ is the
$\mathbf{q} \to 0$ and $\omega \to 0$ limits of the non-interacting
dielectric function $\varepsilon(\mathbf{q}, \omega)$, where
$\mathbf{q}$ is the momentum transfer and $\omega$ is the frequency:
\begin{equation}
  \label{eq:RPA-eps2}
  \varepsilon_{\mathrm{SL}}
  = \lim_{\mathbf{q} \to 0} 1 + \frac{2e^{2}}{\varepsilon_{0} |\mathbf{q}|^{2} \Omega}
  \sum_{\mathrm{k, c, v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
  \left[f(\psi_{\mathrm{c}}) - f(\psi_{\mathrm{v}})\right]
\end{equation}
where $e$ is the unit charge, c, v are the conduction and valence
bands, $E$ is the eigenenergy of individual bands, and $f$ is the
Fermi-Dirac distribution function. Taking the
limit that $L\to\infty$, when
$\varepsilon^{\perp}_{\mathrm{SL}} \approx 1$, we have
$1-1/\varepsilon^{\perp}_{\mathrm{SL}} \approx
(\varepsilon_{\mathrm{SL}}^{\perp} - 1)$. Therefore
$\alpha_{\mathrm{2D}}^{\parallel}$ and $\alpha_{\mathrm{2D}}^{\perp}$
at 0 K can be unified by the same equation:
\begin{equation}
  \label{eq:alpha-RPA}
  \alpha = \frac{2e^{2}}{|q|^{2}A} \sum_{\mathrm{k,c,v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
where the direction is determined by $\mathbf{q}$. Next we will show
that the different behavior of $\psi^{\parallel}$ and $\psi^{\perp}$
give rise to the main text Eqs. \ref{main-eq:2D-Moss-para} and
\ref{main-eq:2D-Moss-perp}.

@@

\subsubsection{Detailed derivations for main text Eq. \ref{main-eq:2D-Moss-perp}}
\label{ssec:theory-2}
For main text Eq. \ref{main-eq:2D-Moss-perp}, treating the in-plane
wave functions as plane wave with form
$\psi^{\parallel}(\rho) \propto e^{i \mathbf{k \rho}}$, the matrix
element of
$<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>$,
when $\mathbf{q}=(0, 0, q_{z})$, becomes\cite{Hybertsen_1987}:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z}
  <\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{qr}}|\psi_{\mathrm{c}}(\mathbf{k+q})>
  &= \frac{1}{A} \int dx \int dy
  e^{i(\mathbf{-k \rho} - \mathbf{q \rho} + \mathbf{(k+q) \rho})}
  \int (\psi^{\perp})^{*}_{\mathrm{v}}(\mathbf{k})e^{-iq_{z}z}\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})\\
  &= <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k+q})>
  \end{aligned}
\end{equation}
Note that the states perpendicular are bound, the integral is
meaningful only when $\mathbf{k=k+q}$ \cite{davies_physics_1997}. By
performing the Taylor expansion of
$e^{-i\mathbf{qr}} \approx 1 - i\mathbf{qr}$, we get:
\begin{equation}
  \begin{aligned}
    \label{eq:matrix-z-2}
    <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|e^{-iq_{z}z}|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
    &\approx <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})> -
    iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>\\
    &= -iq_{z} <\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>
   \end{aligned}
\end{equation}
plug this into Eq. \ref{eq:alpha-RPA} and express the summation over
$k_{x}$ and $k_{y}$ in a continuous form within the Brillouin Zone, we
arrive at:
\begin{equation}
\label{eq:alpha_perp_RPA}
\alpha_{\mathrm{2D}}^{\perp} = \frac{2e^{2}}{(2 \pi) ^{2}} \int d^{2}\mathbf{k}
\sum_{\mathrm{c, v}}
\frac{|<\psi_{\mathrm{v}}(\mathbf{k})|z|\psi_{\mathrm{c}}(\mathbf{k})>|^{2}}
{E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
The formalism is slightly different from Eq.\ref{eq:alpha_para_RPA}.

The out-of-plane wave function $\psi^{\perp}(z)$ is the solution to the Schr\"{o}dinger equation with
Hamiltonian $\mathcal{H} = -\hbar^{2} \nabla^{2}/2m_{e} + V(z)$, where
$\hbar$ is the reduced Planck constant, $m_{e}$ is electron mass and
$V(z)$ is the confined Coulomb potential along the z-direction
created by the nuclei
\cite{davies_physics_1997,ihn_semiconductor_2009}. Although the exact
form for $\psi^{\perp}$ depends on the exact distribution of $V(z)$,
without loss of generality we can assume the electrons are confined in
a potential well of width $\delta$, which is the typical treatment for
semiconductor QWs \cite{ihn_semiconductor_2009}. The allowed bound
states inside the confined region generally has wave vector
$k_{z} \propto n \pi / \delta$. With the total energy
$E_{n}(\mathbf{k}) = {\displaystyle \frac{\hbar^{2} (k_{x}^{2} +
    k_{y}^{2})}{2 m^{\parallel}} + \frac{\hbar^{2} n^{2} \pi^{2}}{2
    m^{\perp} \delta^{2}}}$, where $m^{\parallel}$ and $m^{\perp}$ are
the effective masses parallel and perpendicular to the 2D
plane. Therefore the denominator of Eq. 10 becomes independent of
$\mathbf{k}$, that
$E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k}) =
(n_{\mathrm{c}}^{2} - n_{\mathrm{v}}^{2}) {\displaystyle
  \frac{\hbar^{2} \pi^{2}}{2 m^{\perp} \delta^{2}}}$. On the other
hand, the numerator
$<\psi^{\perp}_{\mathrm{v}}(\mathbf{k})|z|\psi^{\perp}_{\mathrm{c}}(\mathbf{k})>$
is proportional to the confinement length $\delta$ (can be seen using
particle-in-box solution\cite{davies_physics_1997}). In combination,
the individual terms of the summation in the right hand of Eq. 5b is
independent of neither $E_{\mathrm{g}}$ nor \textbf{k}, proving that
$\alpha_{\mathrm{2D}}^{\perp}$ is independent of the band gap.
